Uncovering underlying musical content in the temporal domain \\ \small Leveraging deep learning, inductive bias, and aural music skills to learn embedding spaces from raw waveforms with applications to MIR downstream tasks

### 2023-04-20 introduction/Subsections/Objectives.tex
Updates from Overleaf

In recent years, using neural networks has enabled low-dimensional musical embeddings that capture crucial musical features and allow the computation of content-based similarity between musical passages \cite{Chen2020LearningRecommendation}\cite{KimLEARNINGLOSS}\cite{Hung2022Feature-informedClassification}\cite{WonEmotionStories}. By acquiring an embedding function $f(x) \in \mathbb{R}^d$ that maps a given course $x$ to a vector in the real space $\mathbb{R}^d$, we can measure the similarity between two musical passages using the Euclidean distance, as represented by the formula $D_{i,j} = ||f(x_i) - f(x_j)||^2$. The effectiveness of such systems depends on the level of musical abstraction and the embedding space's ability to interpret music in a way that aligns with human perception. Therefore, a suitable embedding space is crucial to ensure the output aligns with human expectations.

@@

Given a music piece $M$, we aim to learn a single embedding, $e(M)$, simultaneously capturing the underlying content (melody, harmony, critical, holistic elements) and the production aspects. The embedding space should be structured so that similar content and production features are clustered together, effectively isolating the two elements. Formally, our objective is to learn the embedding function $e(M)$ such that:

@@

\subsection{Applications}
%%%%%%%%%%%%%%%%%%%%%%%%
%POTENTIAL APPLICATIONS
%%%%%%%%%%%%%%%%%%%%%%%%%

A self-supervised model that learns an embedding space to retrieve musical content from raw audio signals could have several potential applications and downstream tasks in MIR. Here are some examples:

>>>>

In recent years, using neural networks has enabled low-dimensional musical embeddings that capture crucial musical features and allow the computation of content-based similarity between musical passages \cite{KimLEARNINGLOSS}\cite{Hung2022Feature-informedClassification}. By acquiring an embedding function $f(x) \in \mathbb{R}^d$ that maps a given course $x$ to a vector in the real space $\mathbb{R}^d$, we can measure the similarity between two musical passages using the Euclidean distance, as represented by the formula $D_{i,j} = ||f(x_i) - f(x_j)||^2$. The effectiveness of such systems depends on the level of musical abstraction and the embedding space's ability to interpret music in a way that aligns with human perception. Therefore, a suitable embedding space is crucial to ensure the output aligns with human expectations.

@@

Given a music piece $M$, we aim to learn a single embedding, $e(M)$, simultaneously capturing the underlying content (melody, harmony, critical, holistic elements) and the production aspects. The embedding space should be structured so that similar content and production features are clustered together, effectively isolating the two aspects. Formally, our objective is to learn the embedding function $e(M)$ such that:

@@

\subsection{Academic research and industrial applications}

The advent of self-supervised models that learn embedding spaces for retrieving musical content from audio signals has opened up novel avenues in academic research and industrial applications. By capturing high-level semantic information from raw audio data in an unsupervised manner, these models can potentially revolutionize various aspects of the music industry. Their applications span music information retrieval, sharing learned latent representations for MIR \cite{HamelTransferSimilarity}, recommendation systems, digital rights management, and audio signal processing, fostering interdisciplinary research and innovation in musicology, computer science, and artificial intelligence.

The embedding spaces generated by these models can be utilized to create personalized and intelligent music recommendation systems \cite{Chen2020LearningRecommendation}\cite{epidemic}, facilitate effective enforcement of intellectual property rights, and contribute to developing more sophisticated audio editing-production tools, and technical products in general, \cite{WonEmotionStories}, ultimately enhancing our interaction with music in the digital age.

### 2023-04-20 methods/subsections/model architecture and training strategy.tex
Updates from Overleaf

Contrastive Self-Supervised Learning of Musical Representations (CSSLMR) is an approach that combines the principles of Self-Supervised Learning (SSL) \cite{audioselfsupsurvey} and Contrastive Learning of Musical Representations (CLMR) \cite{CLMR2021} to learn discriminative and valuable representations of music without relying on explicit labels. By leveraging the structure and content of music and contrasting different augmentations or transformations of the same musical piece against randomly (or not) selected pieces, CSSLMR models can effectively capture the underlying structure and known invariance in the data.

@@

\begin{enumerate}
\item \textbf{Data Augmentation:} Create different augmentations or transformations of the same musical piece as positive pairs and use other randomly selected pieces as antagonistic pairs.
\item \textbf{Contrastive Learning:} Train a neural network to minimize a contrastive loss function, encouraging the model to produce similar representations for positive and dissimilar representations for antagonistic pairs.
\item\textbf{Latent Representation:} Utilize the learned representations for various MIR downstream tasks, such as music genre classification or similarity, both in the symbolic and time domains.
\end{enumerate}

@@

The SampleCNN model takes in 1D input data with a single channel. The first layer is a 1D convolutional layer with a kernel size of 3, a stride of 3, and 128 output channels. It is followed by batch normalization and ReLU activation.

After the initial layer, there are nine hidden layers with varying kernel sizes, strides, and channels. Each hidden layer consists of:

\begin{itemize}
    \item A 1D convolutional layer with the specified number of input and output channels, kernel size, a stride of 1, and padding of 1.
    \item Batch normalization.
    \item ReLU activation.
    \item Max pooling with kernel size and stride equal to the stride value of the current layer.
\end{itemize}

@@

\subsection{Audio augmentation and transformation pipeline}
\subsubsection{Positive sample generation}
Let $W$ be the input waveform and $W_p$ be the generated positive waveform. The pipeline applies a series of audio effects and adds noise with a random SNR. Define the following random variables for the effects:

\begin{alignat*}{2}
&gain \sim \text{Uniform}(-12, 0), \quad
&pitch \sim \text{Uniform}(-1200, 1200), \\
&reverb\_params \sim \text{Uniform}(0, 100)^3, \quad
&chorus\_params = \begin{cases}
\text{Uniform}(0.1, 1.0), & i=1,2, \\
\text{Uniform}(20, 55), & i=3, \\
\text{Uniform}(0.1, 0.9), & i=4, \\
\text{Uniform}(0.1, 2.0), & i=5, \\
\text{Uniform}(2, 5), & i=6, \\
\text{Categorical}(\{-s, -t\}), & i=7,
\end{cases}, \\
&drive \sim \text{Uniform}(0, 30), \quad
&stretch \sim \text{Uniform}(0.9, 1.1), \\
&speed \sim \text{Uniform}(0.9, 1.1), \quad
&tremolo\_speed \sim \text{Uniform}(0.1, 100), \\
&tremolo\_depth \sim \text{Uniform}(1, 101), \quad
&snr\_range \sim \text{Uniform}(12, 100).
\end{alignat*}

For white noise, let $N$ be the noise matrix with dimensions (channels, samples) as $W$. Calculate signal power $P_s$ and noise power $P_n$:

\begin{align*}
&P_s = \sum (W_{ij})^2, \
&P_n = \sum (N_{ij})^2.
\end{align*}

Calculate the scaling factor $\alpha$ for noise based on the desired SNR:

\[
\alpha = \sqrt{\frac{P_s}{P_n \times 10^{\frac{snr\_range}{10}}}}
\]

@@

\begin{align*}
&N_s = \alpha \times N, \
&W_n = W + N_s.
\end{align*}

@@

Let $a$ be the original anchor clip and $n$ be the negative clip to be generated. The steps for generating the negative clip $n$ are as follows:

@@

$L_a = \text{shape}(a)$, $\quad T_a = \frac{L_a}{f_s}$,

@@

\item Determine the number of chunks $n_c$ to split the anchor clip into based on a minimum chunk duration $T_c$:
$$
n_c = \left\lfloor\frac{T_a}{T_c}\right\rfloor.
$$

\item Calculate the minimum and maximum chunk lengths in samples:
$$
L_{c,\text{min}} = T_c \times f_s, \quad L_{c,\text{max}} = T_{\text{max}} \times f_s,
$$
where $T_{\text{max}}$ is the maximum chunk duration specified in seconds.

\item Generate $n_c$ chunk lengths $L_{c,1},\ldots,L_{c,n_c}$ as follows:
$$
L_{c,i} = \begin{cases}
L_{c,\text{min}}, & i=1, \\
\text{randint}(L_{c,\text{min}}, L_{c,\text{max}}), & 1<i<n_c, \\
L_a - \sum_{j=1}^{i-1} L_{c,j}, & i=n_c.
\end{cases}
$$

\item Split the anchor clip into chunks $c_1,\ldots,c_{n_c}$ based on the lengths $L_{c,1},\ldots,L_{c,n_c}$:
$$
c_i = a[..., s_i : s_i + L_{c,i}], \quad S_c = \text{cumsum}([0] + [L_{c,1},\ldots,L_{c,n_c-1}]), \quad s_i = S_c[i], \quad i=1,\ldots,n_c.
$$

\item Shuffle the resulting list of chunks randomly to obtain the reordered chunks $c'_1,\ldots,c'_{n_c}$:
$$
c' = \text{shuffle}([c_1,\ldots,c_{n_c}]).
$$

\item Concatenate the shuffled chunks $c'_1,\ldots,c'_{n_c}$ to create the negative clip $n$:
$$
n = \text{concatenate}(c').
$$

>>>>

Contrastive Self-Supervised Learning of Musical Representations (CoSSLeMuR) is an approach that combines the principles of Self-Supervised Learning (SSL) \cite{audioselfsupsurvey} and Contrastive Learning of Musical Representations (CLMR) \cite{CLMR2021} to learn discriminative and valuable representations of music without relying on explicit labels. By leveraging the structure and content of music and contrasting different augmentations or transformations of the same musical piece against randomly (or not) selected pieces, CSSLMR models can effectively capture the underlying structure and known invariance in the data.

@@

\begin{enumerate}
\item \textbf{Data Augmentation:} Create different augmentations or transformations of the same musical piece as positive pairs and use other randomly selected pieces as antagonistic pairs.
\item \textbf{Contrastive Learning:} Train a neural network to minimize a contrastive loss function, encouraging the model to produce similar positive and dissimilar representations for antagonistic pairs.
\item\textbf{Latent Representation:} Utilize the learned representations for various MIR downstream tasks, such as music genre classification or similarity, both in the symbolic and time domains.
\end{enumerate}

@@

The SampleCNN model \cite{CLMR2021} takes in 1D input data with a single channel. The first layer is a 1D convolutional layer with a kernel size of 3, a stride of 3, and 128 output channels. It is followed by batch normalization and ReLU activation.

After the initial layer, there are nine hidden layers with varying kernel sizes, strides, and channels. Each hidden layer consists of the following:

\begin{itemize}
    \item A 1D convolutional layer with the specified number of input and output channels, kernel size, a stride of 1, and padding of 1.

    \begin{equation}
y_{c_o}(n) = \sum_{c_i=1}^{C_{\text{in}}} \sum_{k=-\lfloor K/2 \rfloor}^{\lfloor K/2 \rfloor} x_{c_i}(n - k) \cdot w_{c_o, c_i}(k)
\end{equation}

where:

$y_{c_o}(n)$ is the output for the $c_o$-th output channel at position $n$
$C_{\text{in}}$ is the number of input channels
$x_{c_i}(n)$ is the input for the $c_i$-th input channel at position $n$
$w_{c_o, c_i}(k)$ is the kernel weight for the $c_o$-th output channel and $c_i$-th input channel at position $k$
$K$ is the kernel size
The stride is set to 1 and padding is set to 1 (equal padding $\lfloor K/2 \rfloor$ on both sides)
s
    \item Batch normalization. We process a batch of data containing anchor, positive, and negative samples. It pads the waveforms to the same length, performs online triplet mining to find the hardest negatives, and returns the padded anchors, positives, and hardest negatives.

    Let $A$, $P$, and $N$ represent the anchor, positive, and negative samples, respectively. The maximum length in the batch is given by:

    \begin{equation}
L_{\text{max}} = \max_{\text{item} \in \text{batch}}(\max(\text{length}(A_{\text{item}}), \text{length}(P_{\text{item}}), \text{length}(N_{\text{item}})))
\end{equation}

    The anchor-positive distance and the anchor-negative distance are computed as the Euclidean distance between the corresponding samples:

    \begin{equation}
D_{\text{AP}} = \sqrt{\sum_{i} (A_i - P_i)^2}
\end{equation}

    The hardest negative index for each anchor-positive pair is determined by maximizing the difference between the anchor-negative and anchor-positive distances:

    \begin{equation}
D_{\text{AN}} = \sqrt{\sum_{i} (A_i - N_i)^2}
\end{equation}

    \item ReLU activation. \input{figures/equations/ReLU}

    \item Max pooling with kernel size and stride equal to the stride value of the current layer.
\end{itemize}

@@

We have changed and adapted the original architecture for our experiments to our purpose and needs—the weights initialization and the final average pooling step.

\subsection{Audio augmentation and transformation pipeline}
\subsubsection{Positive sample generation}
Let $W$ be the input waveform and $W_p$ be the generated positive waveform. The pipeline applies audio effects and adds noise with a random SNR. Define the following random variables for the effects:

\begin{equation}
gain \sim \text{Uniform}(-12, 0)
\end{equation}
\begin{equation}
pitch \sim \text{Uniform}(-1200, 1200)
\end{equation}
\begin{equation}
reverb\_params \sim \text{Uniform}(0, 100)^3
\end{equation}
\begin{equation}
chorus\_params = \begin{cases}
\text{Uniform}(0.1, 1.0), & i=1,2, \\
\text{Uniform}(20, 55), & i=3, \\
\text{Uniform}(0.1, 0.9), & i=4, \\
\text{Uniform}(0.1, 2.0), & i=5, \\
\text{Uniform}(2, 5), & i=6, \\
\text{Categorical}(\{-s, -t\}), & i=7,
\end{cases}
\end{equation}
\begin{equation}
drive \sim \text{Uniform}(0, 30)
\end{equation}
\begin{equation}
stretch \sim \text{Uniform}(0.9, 1.1)
\end{equation}
\begin{equation}
speed \sim \text{Uniform}(0.9, 1.1)
\end{equation}
\begin{equation}
tremolo\_speed \sim \text{Uniform}(0.1, 100)
\end{equation}
\begin{equation}
tremolo\_depth \sim \text{Uniform}(1, 101)
\end{equation}
\begin{equation}
snr\_range \sim \text{Uniform}(12, 100)
\end{equation}

Let $N$ be the noise matrix with dimensions [channels, time domain samples] as $W$ for white noise. Calculate signal power $P_s$ and noise power $P_n$:

\begin{equation}
P_s = \sum (W_{ij})^2
\end{equation}
\begin{equation}
P_n = \sum (N_{ij})^2
\end{equation}

Calculate the scaling factor $\alpha$ for noise based on the desired signal-to-noise ratio (SNR):

\begin{equation}
\alpha = \sqrt{\frac{P_s}{P_n \times 10^{\frac{snr\_range}{10}}}}
\end{equation}

@@

\begin{equation}
N_s = \alpha \times N
\end{equation}
\begin{equation}
W_n = W + N_s
\end{equation}

@@

Let $a$ be the original anchor clip and $n$ be the negative clip generated. The steps for generating the negative clip $n$ are as follows:

@@

\begin{equation}
L_a = \text{shape}(a)
\end{equation}
\begin{equation}
T_a = \frac{L_a}{f_s}
\end{equation}

@@

\item Determine the number of chunks $n_c$ to split the anchor clip into based on a minimum chunk duration $T_c$:

\begin{equation}
n_c = \left\lfloor\frac{T_a}{T_c}\right\rfloor
\end{equation}

\item Calculate the minimum and maximum chunk lengths in samples:

\begin{equation}
L_{c,\text{min}} = T_c \times f_s
\end{equation}
\begin{equation}
L_{c,\text{max}} = T_{\text{max}} \times f_s
\end{equation}

where $T_{\text{max}}$ is the maximum chunk duration specified in seconds.

\item Generate $n_c$ chunk lengths $L_{c,1},\ldots,L_{c,n_c}$ as follows:

\begin{equation}
L_{c,i} = \begin{cases}
L_{c,\text{min}}, & i=1, \\
\text{randint}(L_{c,\text{min}}, L_{c,\text{max}}), & 1<i<n_c, \\
L_a - \sum_{j=1}^{i-1} L_{c,j}, & i=n_c.
\end{cases}
\end{equation}

\item Split the anchor clip into chunks $c_1,\ldots,c_{n_c}$ based on the lengths $L_{c,1},\ldots,L_{c,n_c}$:

\begin{equation}
c_i = a[..., s_i : s_i + L_{c,i}]
\end{equation}
\begin{equation}
S_c = \text{cumsum}([0] + [L_{c,1},\ldots,L_{c,n_c-1}])
\end{equation}
\begin{equation}
s_i = S_c[i]
\end{equation}
\begin{equation}
i=1,\ldots,n_c
\end{equation}

\item Shuffle the resulting list of chunks randomly to obtain the reordered chunks $c'_1,\ldots,c'_{n_c}$:

\begin{equation}
c' = \text{shuffle}([c_1,\ldots,c_{n_c}])
\end{equation}

\item Concatenate the shuffled chunks $c'_1,\ldots,c'_{n_c}$ to create the negative clip $n$:

\begin{equation}
n = \text{concatenate}(c')
\end{equation}
